Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-09-16 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

A quadrant photo diode (QPD) is a frequently used position detection 
system for optical tweezers due to their high sampling rate, high 
degree of precision, and ease of set up. The QPD is constructed of 
four photo diodes assembled in a quadrant formation, when a particle 
is trapped the interference pattern produced is focused onto the QPD, 
with the maximum intensity mapping to the particle's centre of mass. 
By summing the voltages of the horizontal and vertical quadrants together 
the particle's centre of mass is tracked in the x-y plane. Axial 
displacement can be estimated by observing the change in the total 
voltage of the QPD. The outputted signal gives an indication of the 
particle's relative displacement from the beam focus, but in order to 
convert the signal to distance units the trap needs to be calibrated 
(assuming a linear response curve).
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
	\caption{Comparison between QPD and Lateral effect photodiodes.
	The four quadrants of a QPD (left) experience different photocurrents
	based on the total intensity of light incident on each section 
	(labelled A, B, C, D). 
	Whereas a Lateral effect sensor (right) uses the resistive properties
	of the photodiode surface to vary the create different photocurrents 
	passing through the anodes A, B, C, and D.}
\end{figure}

@@

\subsection{Stationary beam}
\label{sec:stationary}
Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
event in supersaturated glycine solution ($S=1.03$), the beam is a stationary being 
$\approx3.5 \mu m$ from the droplet edge. After a period of roughly 5 minutes a 
nucleus forms at the trap focus, growing quickly (growth rate was approximated using
imageJ to be on the order of $700\ \mu m^2/min$) from the focal point of the trap 
until after roughly $6$ seconds the crystal escapes. Comparing to previous literature
using optical tweezers shows that the growth rate is only loosely connected to the 
solutions supersaturation. Local conditions play a much larger role in the growth 
rate than just the concentration \cite{Flannigan2023}. A likely reason that the trap 
is escaped is due to the fact that crystal is far too large to be held in place and 
is in fact still growing as the solution is supersaturated. The key take away to 
remember is that the beam has no influence over the crystal shape, instead it 
grows outward from the trap focus. Furthermore, due to the fact that the solution
is supersaturated the crystal growth cannot be contained to the trap focus. Instead 
the crystal escapes as its size exceeds the trap focus. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.63\linewidth]{frames_no_beam_movement.pdf}
	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape 
	the trap.}
	\label{fig:stationary_beam}
\end{figure}

\subsection{Moving Beam}
\label{sec:moving}
To test if a rapidly moving silica bead could generate the necessary shear rate 
for crystal nucleation we wanted to see if a trapped silica bead could be trapped
and moved in an aqueous solution. $20\ \mu L$ of glycine and water ($S=1.03$) was 
added to $10\ \mu L$ of a dilute water-silica mixture making the solution 
unsaturated ($S\approx0.7$). However, due to the beam's motion we instead 
encountered several unexpected results.

Shown below in Fig.~\ref{fig:eliptical_beam_1} where we have the laser focus 
moving in a small elliptical pattern; interestingly. After a few minutes we 
noted the appearance of several small droplets (see figure~\ref{fig:cluster_trapping}
for a larger example of these droplets). These appear too have a wide distribution
of sizes, unlike silica microspheres which have a uniform radius of $1.57\ 
\mu m$. We surmise these could be small clusters of Glycine that had previously 
been shown to form when the aqueous solutions where irradiated with a 
focused laser \cite{Tsuboi2009, Gowayed2021}. While no droplets are seen 
directly entering the focus a nucleus forms close to the droplet edge, 
unlike in Fig.~\ref{fig:stationary_beam} the crystal does not grow out from 
the focal point evenly. Due to the galvano mirror, the crystal is simultaneously 
being moved by and growing around the focal point of the trap. Because of 
this the crystal nucleus lacks a clear morphology at first. Until roughly 
$20\ s$ the crystal reaches a almost prismatic structure, with further 
irradiation increasing the size.

@@

Nucleation in undersaturated conditions has been reported previously in $D_2O$ 
\cite{Rungsimanon2010} and $H_2O$ \cite{Flannigan2023}, though not involving a 
moving beam. This modification allows for the crystal growth to be localised to
a specific region of the bulk crystal whereas with a stationary beam there is no
control over the crystal morphology. In fact this allows for a much finer control
over the exact shape of the crystal nucleus, in some cases allowing for growth out
of the viewing plane as shown in

\begin{figure}
	\centering
\end{figure}

@@

\section{Influence of a moving beam front on seed crystals}
\textcolor{red}{Need to write up these results. To summarise the crystal 
	dissolved around the laser focus instead of growing as the seed crystals 
	are in the solution and not close to the air-solution interface}

>>>>

A quadrant photo diode (QPD) is a frequently used position detection 
system for optical tweezers due to their high sampling rate, high 
degree of precision, and ease of set up. The QPD is constructed of 
four photo diodes assembled in a quadrant formation, when a particle 
is trapped the interference pattern produced is focused onto the QPD, 
with the maximum intensity mapping to the particle's centre of mass. 
By summing the voltages of the horizontal and vertical quadrants together 
the particle's centre of mass is tracked in the x-y plane. Axial 
displacement can be estimated by observing the change in the total 
voltage of the QPD. The outputted signal gives an indication of the 
particle's relative displacement from the beam focus, but in order to 
convert the signal to distance units the trap needs to be calibrated 
(assuming a linear response curve).
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
	\caption{Comparison between QPD and Lateral effect photodiodes.
	The four quadrants of a QPD (left) experience different photocurrents
	based on the total intensity of light incident on each section 
	(labelled A, B, C, D). 
	Whereas a Lateral effect sensor (right) uses the resistive properties
	of the photodiode surface to vary the create different photocurrents 
	passing through the anodes A, B, C, and D.}
	\label{fig:position_detectors}
\end{figure}

@@

\subsection{Stationary beam}
\label{sec:stationary}
As an example case, consider below in Fig.~\ref{fig:stationary_beam} 
the frames taken from a nucleation event in supersaturated glycine 
solution ($S=1.03$), the beam is a stationary being $\approx3.5 
\mu m$ from the droplet edge. After a period of roughly 5 minutes a 
nucleus forms at the trap focus, growing quickly from the focal point 
of the trap until after roughly $6$ seconds the crystal escapes. 
The growth rate was approximated using imageJ to be on the order of 
$700\ \mu m^2/min$. Comparing to previous literature using optical 
tweezers shows that the growth rate is only loosely connected to the 
solutions supersaturation. Local conditions play a much larger role 
in the growth rate than just the concentration \cite{Flannigan2023}. 
A likely reason that the trap is escaped is due to the fact that 
crystal is far too large to be held in place and is in fact still 
growing as the solution is supersaturated. The key take away to 
remember is that the beam has no influence over the crystal shape, 
instead it grows outward from the trap focus. Furthermore, due to 
the fact that the solution is supersaturated the crystal growth 
cannot be contained to the trap focus. Instead the crystal 
escapes as its size exceeds the trap focus. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.63\linewidth]{frames_no_beam_movement.pdf}
	\caption{Laser induced nucleation at the edge of a droplet of supersaturated glycine solution. (b) shows the first instance 
	of a crystal nucleus, growing quickly through (c)-(e) until 
	after $6.665\ s$ the crystal begins to escape the trap.}
	\label{fig:stationary_beam}
\end{figure}

Fig.~\ref{fig:stationary_beam} makes it clear that in general, the
trap focus can be used to induce nucleation. Previous reports
have shown that this can occur over a range of supersaturations 
\cite{Rungsimanon2010,Liao2022,Flannigan2023}. In every case they
utilise a stationary beam as this is often easier to track. Using
a galvno mirror we now consider the impact of a moving beam.

\subsection{Moving Beam}
\label{sec:moving}
To test if a rapidly moving silica bead could generate the 
necessary shear rate for crystal nucleation, a silica bead 
be trapped in an aqueous solution. $20\ \mu L$ of glycine 
and water ($S=1.03$) was added to $10\ \mu L$ of a dilute 
water-silica mixture making the solution unsaturated ($S
\approx0.7$). While trying to bring a silica bead towards
the droplet edge several smaller droplets were seen within
the fluid (see figure~\ref{fig:cluster_trapping} for a 
larger example of these droplets).

These appear too have a wide distribution of sizes, unlike 
silica microspheres which have a uniform radius of $1.57\ 
\mu m$. We surmise these could be small clusters of Glycine 
that had previously been shown to form when the aqueous 
solutions where irradiated with a focused laser \cite{Tsuboi2009, 
Gowayed2021}. While no droplets are seen directly entering 
the focus a nucleus forms close to the droplet edge, unlike 
in Fig.~\ref{fig:stationary_beam} the crystal does not grow 
out from the focal point evenly. Due to the galvano mirror, 
the crystal is simultaneously being moved by and growing 
around the focal point of the trap. Because of this the 
crystal nucleus lacks a clear morphology at first. This is 
in contrast to previous works with undersaturated solutions, 
where even then the laser focus has little to no impact on 
the crystal form \cite{Rungsimanon2010}.

@@

Nucleation in undersaturated conditions has been reported 
previously in $D_2O$ \cite{Rungsimanon2010} and $H_2O$ 
\cite{Flannigan2023}, though not involving a moving beam. 
This modification allows for the crystal growth to be 
localised to a specific region of the bulk crystal whereas 
with a stationary beam there is no control over the crystal 
morphology. In fact this allows for a much finer control
over the exact shape of the crystal nucleus, in some cases 
allowing for growth out of the viewing plane as shown in
fig.XX

@@

\section{Influence of a moving beam front on seed crystals}
Localised crystal growth could have useful applications for 
tailoring seed crystals. To understand how the a moving beam
front could be

### 2024-09-16 Latex_files/Thesis/chapters/chapter 5/main.tex
updates

\chapter{Detection and Characterisation of rotational spherical aggregate rotational dynamics}
\label{chapter:simulated_detection}
As outlined in the end of Chapter 4, one of the difficulties in characterising 
interactions with asymmetric objects is the coupled motion between translation 
and rotation. This can be somewhat ignored if the rotational motion is periodic,
like when the power spectra of elliptical polystyrene particles was fitted by 
Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, however in the case where 
rotational motion is stochastic the problem is more complex. For example, when a
novel trap characterisation technique was implemented by Saffron \textit{et al}
\cite{BarZiv1997, Meller1998}, they were able to use dynamic light scattering to
characterise both the axial and lateral trap stiffness acting on microspheres. The
only drawback admitted to in their work was that the technique was constrained to
isotropic scatters as their theoretical model for describing the auto-correlation
function was predicated on the fact that any variations in the signal are due to
the particles translational motion within the confines of a cylindrical trap 
\cite{BarZiv1997}. However, it is somewhat naive to assume that the orientational 
change

\section{Simulative QPD}
A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.

\section{Monitoring Stochastic rotational motion using static light scattering}
Orientational dynamics to an anisotropically scattering shape are difficult to 
characterise due to the coupling of rotational and translational effects. As 
shown in the previous chapter the dynamics of even simple dimers are heavily 
dependent on the particle's position and orientation, this is reflected in 
literature where several engineering solutions have been devised to decouple 
translational and orientational motion \textcolor{red}{[ , ]}. While a majority 
of latter work has been focused on nano-particles (falling into the Rayleigh 
Regime) and utilising florescence

\subsection{Coordinate System}
In the case of our probe beam the +z direction points from the surface of the 
probe directly to the centre of diffusion in our dimer, with the +y direction 
pointing towards what would be our trapping beam, and the +x direction pointing 
into the page. The origin of our coordinate system is fixed on the centre of 
the focus of our trapping beam, meaning as our dimer's centre of diffusion moves 
the origin is kept constant.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:Conclusion}

>>>>

\chapter{Detection and Characterisation of rotational spherical aggregate rotational dynamics}
\label{chapter:simulated_detection}
As outlined in the end of Chapter 4, one of the difficulties 
in characterising interactions with asymmetric objects is 
the coupled motion between translation and rotation. In order
to characterise the optical trap a position detection system 
such as a lateral effect photodiode or a quadrant photo diode
(see fig.\ref{fig:position_detectors}). Typically, a position 
detection system assumes that there exists a linear relationship 
between the a particle's displacement and the detected signal.
However, this is not always the case as demonstrated in chapter
4; as dimer's demonstrate non-harmonic trapping beyond a certain 
displacement. In addition, the rotational effects of the trapped
dimer become a more significant factor when characterising the 
trap strength. This can have unintended effects when it comes to
areas of research that require a precise understanding of the 
hydrodynamic behaviour of particles. As such there has been a 
concerted effort to either eliminate rotational motion, or describe
how the rotational behaviour influences the trajectory reported 
by the position detector.

Rotational motion about a single axis is easiest to account for.  
When the power spectra of elliptical polystyrene particles was 
fitted by Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, they 
assumed that the rotational motion was purely in the transverse
plane. As such they did not have to account for any variance in 
the trapping strength due to orientation nor did they need to
consider non-periodic rotational behaviour. In the case where 
rotational motion is stochastic the problem is more complex. 
For example, when an optical fibre trap characterisation 
technique was implemented by Saffron \textit{et al} 
\cite{BarZiv1997, Meller1998}, they were able to use dynamic 
light scattering to characterise both the axial and lateral 
trap stiffness acting on microspheres. The only drawback 
admitted to in their work was that the technique was constrained 
to isotropic scatters as their theoretical model for describing 
the auto-correlation function was predicated on the fact that 
any variations in the signal are due to the particles 
translational motion within the confines of a cylindrical trap 
\cite{BarZiv1997}. Where the upper limit of the cylindrical 
trap is given by the Rayleigh range ($z_R = n\omega_0/NA$).
However, as demonstrated by the results from Chapter 4.1, 
the axial traps of spherical aggregates is often situated 
far beyond the Rayleigh range (for a 1.2 NA laser this is 
$\pm5.985 \mu m$).

To accurately describe the full dynamics of an anisotropic
scatterer its clear that additional characterisation 
techniques are necessary. In this chapter we demonstrate 
two such techniques, quadrant photo-diode simulation, and 
an optical fibre static light scattering arrangement. The 
former is based on the work of Rohrbach and Stelzer 
\cite{Rohrbach2002}, where their work is limited to single 
spheres, we extend the possible particle space to allow for 
any theoretical combination of spheres. The results from
chapter 4 demonstrate that the rotational motion causes 
significant deviations between the computed trapping strength
and the expected trapping strength according to Lorenz-Mie
theory. By utilizing vector regression we can accurately 
recreate a particle's trajectory simply from the QPD signal, 
both in the transverse plane and along the beam axis.

The latter method builds upon \cite{BarZiv1997} but now 
accounts for the full rotational motion. This was initially
developed in tandem with the work conducted in chapter 3,
with the idea being that if a crystal nucleus formed on a 
rotating sphere how would that nucleus influence the motion
of the rotating sphere. Using static light scattering we can 
map the outputted signal from a optically trapped dimer to 
its expected orientation in real time. Thus allowing us to 
estimate the optical torque being applied to the dimer by 
the trapping beam. As demonstrated in chapter 4, the 
influence of a circularly polarised beam can result in 
gyroscopic precession for specifically sized dimers.

We start by outlining the working principle behind the
optical fibre detection system, before demonstrating 
how it can be used to instantaneously predict 
orientational information on a trapped particle. We then
consider common factors of error in the characterisation
process, such as signal error, and incorrect particle 
sizing. Both of which have a significant impact on the 
performance of the model, but by utilising Bayesian 
inference, and time averaging we can better refine our 
model.

\section{Monitoring Stochastic rotational motion using static light scattering}
Part of the optical fibre arrangement borrows code from the 
optical tweezer simulations discussed in Chapter 2 and 4. 
Rather than repeat the simulation details we will instead 
only focus on the specifics of 
\subsection{Coordinate System}
In the case of our probe beam the +z direction points from the surface of the 
probe directly to the centre of diffusion in our dimer, with the +y direction 
pointing towards what would be our trapping beam, and the +x direction pointing 
into the page. The origin of our coordinate system is fixed on the centre of 
the focus of our trapping beam, meaning as our dimer's centre of diffusion moves 
the origin is kept constant.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{QPD Simulations}
A QPD is simply a measure of the total electric field incident on a photo diode, in order to accurately simulate a QPD response careful consideration of how the Electromagnetic fields are defined is required.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:Conclusion}
